---
title: "SIR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SIR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(SIRModel)
```

To begin the SIR model simulation, we must first start with a generic matrix to represent each cell.
We can create any size matrix and start any specific location as infected cells.
Below is an example of a matrix with only the corners starting as infected cells.

```{r}
x <- create_crnr_matrix(5,5)
```
The "1's" in the matrix represent infected cells.
The "0's" in the matrix represent healthy cells that can be infected.

A single time step can be simulated using the following function.
Each step requires an infection matrix as input and the probability that the infection will spread to other cells.

```{r}
infection_step(x, 0.5)
```
The "2's" represent previously infected cells that can no longer spread or get infected. (Dead cells)

You can run the simulation until there are no infected cells remaining in the given input matrix using the simulation function.
The arguments in order:
  Probability of infection spreading
  Rows in matrix
  Columns in matrix
  Probability of cells starting as infected
```{r}
simulate_sir(0.5, 10, 10, 0.125)
```

The output will produce a list of how many steps the simulation took, the starting infection probability, the probability of infection spreading, the final proportion of cells that became infected, and the final infection matrix.

"0's" represent health cells.
"1's" represent infected cells.
"2's" represent dead cells.

You can also run simulations with using a sequence of different probabilities. The seqeunce starts at 0.1 and ends at 0.9, the second argument specifies what value the sequence should step by. In this case we are using a step value of 0.1.
```{r}
simulation(0.5, 0.1, 10, 10)
```
The output will give you a DataFrame of each probability tested, the number of steps the simulation took before completing for each respective probability, and the proportion of total cells that were infected at the end of the simulation.

Similarly, you can run a single simulation multiple times to see which cells are most likely to become infected.
```{r}
multiple_sims(0.5, 1000, 15, 15)
```

This output generates a heat map of the cells that are most likely to be infected by summing the matrices for the given number of runs you specified.
On this graph, the darker the cells are, the more likely they are to become infected.
