---
title: "SIRSimulation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SIRSimulation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(SIRSimulation)
```

## Create an initial matrix

SIRSimulation uses `0` for susceptible cells and `1` for infected cells in the
initial population matrix.

```{r initial-matrix}
initial <- create_matrix(row = 10, col = 10, start_inf = 0.1, seed = 94128)
initial
```

Corner- and center-seeded matrices are also available:

```{r seeded-matrices}
create_crnr_matrix(row = 10, col = 10)
create_cntr_matrix(row = 10, col = 10)
```

## Run one SIR simulation

```{r one-run}
result <- simulate_sir(prob_infect = 0.25, input_matrix = initial, seed = 94128)
result$steps
result$inf_prop
result$final_matrix
```

## Inspect history

Each simulation records susceptible, infected, and recovered counts by step.

```{r history}
result$history
```

## Run repeated simulations

```{r repeated-runs}
simulate_many_runs(prob_infect = 0.25, input_matrix = initial, runs = 10, seed = 94128)
```

## Sweep infection probabilities

```{r probability-sweep}
simulate_inf_seq(input_matrix = initial, step = 0.1, seed = 94128)
```

## Create a recovered-cell heatmap

`multiple_run_heatmap()` returns a matrix of recovered counts by cell. Set
`plot = TRUE` to draw the heatmap with base R graphics.

```{r heatmap}
heatmap_counts <- multiple_run_heatmap(
  prob_infect = 0.25,
  input_matrix = initial,
  runs = 10,
  plot = FALSE,
  seed = 94128
)
heatmap_counts
```
