---
title: "SIRSimulation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SIRSimulation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(SIRSimulation)
```

## State encoding

Population matrices use simple numeric states:

- `0`: susceptible
- `1`: infected
- `2`: recovered
- `3`: deceased

Initial matrices usually contain only susceptible and infected cells.

```{r initial-matrix}
initial <- create_matrix(row = 10, col = 10, start_inf = 0.1, seed = 94128)
initial
```

Corner- and center-seeded matrices are also available:

```{r seeded-matrices}
create_crnr_matrix(row = 10, col = 10)
create_cntr_matrix(row = 10, col = 10)
```

## Models

`model = "SIR"` makes infected cells recover after their infectious step.

```{r sir-run}
sir_result <- simulate_sir(
  prob_infect = 0.25,
  input_matrix = initial,
  model = "SIR",
  seed = 94128
)
sir_result$steps
sir_result$inf_prop
sir_result$final_matrix
```

`model = "SIS"` makes infected cells return to susceptible.

```{r sis-run}
simulate_sir(
  prob_infect = 0,
  input_matrix = matrix(1, nrow = 1, ncol = 1),
  model = "SIS",
  seed = 94128
)
```

`model = "SIRS"` uses `imm_prob` to decide whether infected cells recover or
return to susceptible.

```{r sirs-run}
simulate_sir(
  prob_infect = 0,
  input_matrix = matrix(1, nrow = 1, ncol = 1),
  model = "SIRS",
  imm_prob = 0.7,
  seed = 94128
)
```

## Mortality

Set `allow_death = TRUE` to let infected cells become deceased with probability
`fat_prob`. Cells that do not die follow the active model transition.

```{r mortality-run}
simulate_sir(
  prob_infect = 0,
  input_matrix = matrix(1, nrow = 1, ncol = 1),
  model = "SIR",
  allow_death = TRUE,
  fat_prob = 1,
  seed = 94128
)
```

## History

Each simulation records susceptible, infected, and recovered counts by step.

```{r history}
sir_result$history
```

## Repeated runs

```{r repeated-runs}
simulate_many_runs(
  prob_infect = 0.25,
  input_matrix = initial,
  runs = 10,
  model = "SIR",
  seed = 94128
)
```

## Probability sweeps

```{r probability-sweep}
simulate_inf_seq(
  input_matrix = initial,
  step = 0.1,
  model = "SIR",
  seed = 94128
)
```

## Heatmaps

`multiple_run_heatmap()` returns a matrix of recovered counts by cell. Set
`plot = TRUE` to draw the heatmap with base R graphics.

```{r heatmap}
heatmap_counts <- multiple_run_heatmap(
  prob_infect = 0.25,
  input_matrix = initial,
  runs = 10,
  model = "SIR",
  plot = FALSE,
  seed = 94128
)
heatmap_counts
```
