---
title: "SIRSsim"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SIRSsim}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(SIRSsim)
```

## State encoding

Population matrices use simple numeric states:

- `0`: susceptible
- `1`: infected
- `2`: recovered
- `3`: deceased

Initial matrices usually contain only susceptible and infected cells.
`create_random_matrix()` infects an exact number of randomly selected cells.

```{r initial-matrix}
initial <- create_random_matrix(
  row = 10,
  col = 10,
  start_infected = 8,
  seed = 94128
)
initial
```

Use `start_infected = 0` to create an all-susceptible starting matrix.

```{r initial-count-matrix}
create_random_matrix(
  row = 10,
  col = 10,
  start_infected = 0,
  seed = 94128
)
```

Center- and corner-seeded matrices are also available. These helpers accept the
same parameters as `create_random_matrix()` for a consistent interface, but
their named patterns are fixed: the center helper infects the center cell, and
the corner helper infects all four corners.

```{r seeded-matrices}
create_center_matrix(row = 10, col = 10, start_infected = 1, seed = NULL)
create_corner_matrix(row = 10, col = 10, start_infected = 1, seed = NULL)
```

## Models

`model = "SIR"` makes infected cells recover after their infectious step.

```{r sir-run}
sir_result <- simulate_sir(
  prob_infect = 0.25,
  input_matrix = initial,
  model = "SIR",
  seed = 94128
)
sir_result$steps
sir_result$inf_prop
sir_result$final_matrix
```

`model = "SIS"` makes infected cells return to susceptible.

```{r sis-run}
simulate_sir(
  prob_infect = 0,
  input_matrix = matrix(1, nrow = 1, ncol = 1),
  model = "SIS",
  seed = 94128
)
```

`model = "SIRS"` uses `imm_prob` to decide whether infected cells recover or
return to susceptible.

```{r sirs-run}
simulate_sir(
  prob_infect = 0,
  input_matrix = matrix(1, nrow = 1, ncol = 1),
  model = "SIRS",
  imm_prob = 0.7,
  seed = 94128
)
```

## Mortality

Set `allow_death = TRUE` to let infected cells become deceased with probability
`fat_prob`. Cells that do not die follow the active model transition.

```{r mortality-run}
simulate_sir(
  prob_infect = 0,
  input_matrix = matrix(1, nrow = 1, ncol = 1),
  model = "SIR",
  allow_death = TRUE,
  fat_prob = 1,
  seed = 94128
)
```

## History

Each simulation records susceptible, infected, and recovered counts by step.

```{r history}
sir_result$history
```

## Full logs for survival analysis

Set `full_log = TRUE` to return one row per cell per simulation step. The log
includes step `0`, before any transitions occur.

```{r full-log}
logged_result <- simulate_sir(
  prob_infect = 0.25,
  input_matrix = initial,
  model = "SIR",
  seed = 94128,
  full_log = TRUE
)

head(logged_result$full_log)
```

The full log contains enough information to derive per-cell survival-analysis
fields such as first infection, recovery, death, last observed step, event
indicators, time to event, and final state.

```{r survival-ready}
cell_log <- logged_result$full_log

survival_ready <- do.call(
  rbind,
  lapply(split(cell_log, cell_log$cell_id), function(cell) {
    first_infected_step <- if (any(cell$was_infected)) min(cell$step[cell$was_infected]) else NA
    recovered_step <- if (any(cell$was_immune)) min(cell$step[cell$was_immune]) else NA
    death_step <- if (any(cell$was_deceased)) min(cell$step[cell$was_deceased]) else NA
    last_observed_step <- max(cell$step)

    data.frame(
      cell_id = cell$cell_id[1],
      first_infected_step = first_infected_step,
      recovered_step = recovered_step,
      death_step = death_step,
      last_observed_step = last_observed_step,
      event_infection = !is.na(first_infected_step),
      event_recovery = !is.na(recovered_step),
      event_death = !is.na(death_step),
      time_to_infection = first_infected_step,
      time_to_recovery = recovered_step,
      time_to_death = death_step,
      final_state = cell$state[cell$step == last_observed_step]
    )
  })
)

head(survival_ready)
```

## Repeated runs

```{r repeated-runs}
simulate_many_runs(
  prob_infect = 0.25,
  input_matrix = initial,
  runs = 10,
  model = "SIR",
  seed = 94128
)
```

## Probability sweeps

```{r probability-sweep}
simulate_inf_seq(
  input_matrix = initial,
  step = 0.1,
  model = "SIR",
  seed = 94128
)
```

## Heatmaps

`multiple_run_heatmap()` returns a matrix of infection counts by cell. Each
cell is counted once per run if it was infected at any point in that run,
including cells that later recovered, became susceptible again, or died. Set
`plot = TRUE` to draw the heatmap with base R graphics.

```{r heatmap}
heatmap_counts <- multiple_run_heatmap(
  prob_infect = 0.25,
  input_matrix = initial,
  runs = 10,
  model = "SIR",
  plot = FALSE,
  seed = 94128
)
heatmap_counts
```
