---
title: "SIRSsim"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SIRSsim}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(SIRSsim)
```

## What SIRSsim simulates

SIRSsim represents a population as a matrix, or grid. Each cell in the grid can
be read as one person, place, or unit in the population. At each simulation
step, infected cells can spread infection to susceptible neighbors, then move
to their next state.

The package is intended for teaching, exploration, and code experiments. It is
not a calibrated public-health forecasting model.

## Models in plain English

SIRSsim supports three related model choices:

- `SIR`: susceptible cells can become infected; infected cells then recover and
  stay recovered.
- `SIS`: susceptible cells can become infected; infected cells then become
  susceptible again, so they can be infected again later.
- `SIRS`: susceptible cells can become infected; infected cells recover with
  probability `imm_prob` or otherwise become susceptible again. In this package,
  recovered cells stay recovered once they reach that state.

Optional mortality can be layered onto any model with `allow_death = TRUE`.
When mortality is enabled, infected cells can become deceased with probability
`fat_prob` before following the model's nonfatal transition.

## State encoding

Population matrices use simple numeric states:

- `0`: susceptible
- `1`: infected
- `2`: recovered
- `3`: deceased

Initial matrices usually contain only susceptible and infected cells.
Infection can spread from any of the eight neighboring cells: horizontal,
vertical, or diagonal.

## Starting matrices

`create_random_matrix()` creates a matrix with an exact number of randomly
selected infected cells.

```{r initial-matrix}
initial <- create_random_matrix(
  row = 10,
  col = 10,
  start_infected = 8,
  seed = 94128
)
initial
```

Use `start_infected = 0` to create an all-susceptible starting matrix.

```{r initial-count-matrix}
create_random_matrix(
  row = 10,
  col = 10,
  start_infected = 0,
  seed = 94128
)
```

Center- and corner-seeded matrices are also available. These helpers accept the
same parameters as `create_random_matrix()` for a consistent interface, but
their named patterns are fixed: the center helper infects the center cell, and
the corner helper infects all four corners.

```{r seeded-matrices}
create_center_matrix(row = 10, col = 10, start_infected = 1, seed = NULL)
create_corner_matrix(row = 10, col = 10, start_infected = 1, seed = NULL)
```

## Important parameters

Most probabilities are written as numbers between `0` and `1`, where `0` means
"never" and `1` means "always."

| Concept | Argument | Meaning |
| --- | --- | --- |
| Infection probability | `prob_infect` | Chance that one infected neighbor infects a susceptible cell during a step. Multiple infected neighbors increase the overall infection chance. |
| Immunity probability | `imm_prob` | In `model = "SIRS"`, chance that an infected cell becomes recovered after its infectious step instead of becoming susceptible again. Ignored by `SIR` and `SIS`. |
| Fatality probability | `fat_prob` | Chance that an infected cell becomes deceased when `allow_death = TRUE`. Ignored when mortality is not enabled. |
| Starting matrix | `input_matrix` | The grid of starting states passed to simulation functions. Use the matrix helpers above to create common starting patterns. |
| Seed | `seed` | Optional random seed for reproducible random matrices and simulation paths. |
| Full log | `full_log` | When `TRUE`, `simulate_sir()` includes one row per cell per step in the returned `full_log` data frame. |

## Reproducibility

Random choices occur when cells are selected for a random starting matrix, when
infection spreads, when SIRS immunity is assigned, and when optional mortality
is enabled. Pass the same `seed` with the same inputs to reproduce the same
result.

For repeated-run helpers such as `simulate_many_runs()` and
`multiple_run_heatmap()`, one seed makes the whole sequence of runs
reproducible.

## One simulation

`model = "SIR"` makes infected cells recover after their infectious step.

```{r sir-run}
sir_result <- simulate_sir(
  prob_infect = 0.25,
  input_matrix = initial,
  model = "SIR",
  seed = 94128
)
sir_result$steps
sir_result$inf_prop
sir_result$final_matrix
```

## Reading outputs

`simulate_sir()` returns a list with summary fields and, optionally, a detailed
log.

| Output | Meaning |
| --- | --- |
| `steps` | Number of simulation steps completed before no infected cells remained. Step `0` is the starting matrix and is included in `history` and `full_log`, not in this count. |
| `prob_infect` | Infection probability used for the run. |
| `inf_prop` | Proportion of cells that are recovered in the final matrix. In a basic `SIR` run without death this is often the final outbreak size; in `SIS`, `SIRS`, or mortality scenarios it is not the same as "ever infected." |
| `history` | Data frame with susceptible, infected, and recovered counts at each step, including step `0`. Deceased cells appear in `final_matrix` and `full_log`, but are not a separate `history` column. |
| `final_matrix` | Matrix of final cell states after the simulation stops. |
| `full_log` | Only returned when `full_log = TRUE`; a long data frame with one row per cell per step, including cell position, state, model settings, and state indicators. |

For "ever infected" questions, use `full_log` or
`multiple_run_heatmap()` instead of `inf_prop`.

## SIS and SIRS examples

`model = "SIS"` makes infected cells return to susceptible.

```{r sis-run}
simulate_sir(
  prob_infect = 0,
  input_matrix = matrix(1, nrow = 1, ncol = 1),
  model = "SIS",
  seed = 94128
)
```

`model = "SIRS"` uses `imm_prob` to decide whether infected cells recover or
return to susceptible.

```{r sirs-run}
simulate_sir(
  prob_infect = 0,
  input_matrix = matrix(1, nrow = 1, ncol = 1),
  model = "SIRS",
  imm_prob = 0.7,
  seed = 94128
)
```

## Mortality

Set `allow_death = TRUE` to let infected cells become deceased with probability
`fat_prob`. Cells that do not die follow the active model transition.

```{r mortality-run}
simulate_sir(
  prob_infect = 0,
  input_matrix = matrix(1, nrow = 1, ncol = 1),
  model = "SIR",
  allow_death = TRUE,
  fat_prob = 1,
  seed = 94128
)
```

## History

Each simulation records susceptible, infected, and recovered counts by step.

```{r history}
sir_result$history
```

## Full logs for survival analysis

Set `full_log = TRUE` to return one row per cell per simulation step. The log
includes step `0`, before any transitions occur.

```{r full-log}
logged_result <- simulate_sir(
  prob_infect = 0.25,
  input_matrix = initial,
  model = "SIR",
  seed = 94128,
  full_log = TRUE
)

head(logged_result$full_log)
```

The full log contains enough information to derive per-cell survival-analysis
fields such as first infection, recovery, death, last observed step, event
indicators, time to event, and final state.

```{r survival-ready}
cell_log <- logged_result$full_log

survival_ready <- do.call(
  rbind,
  lapply(split(cell_log, cell_log$cell_id), function(cell) {
    first_infected_step <- if (any(cell$was_infected)) min(cell$step[cell$was_infected]) else NA
    recovered_step <- if (any(cell$was_immune)) min(cell$step[cell$was_immune]) else NA
    death_step <- if (any(cell$was_deceased)) min(cell$step[cell$was_deceased]) else NA
    last_observed_step <- max(cell$step)

    data.frame(
      cell_id = cell$cell_id[1],
      first_infected_step = first_infected_step,
      recovered_step = recovered_step,
      death_step = death_step,
      last_observed_step = last_observed_step,
      event_infection = !is.na(first_infected_step),
      event_recovery = !is.na(recovered_step),
      event_death = !is.na(death_step),
      time_to_infection = first_infected_step,
      time_to_recovery = recovered_step,
      time_to_death = death_step,
      final_state = cell$state[cell$step == last_observed_step]
    )
  })
)

head(survival_ready)
```

## Repeated runs

```{r repeated-runs}
simulate_many_runs(
  prob_infect = 0.25,
  input_matrix = initial,
  runs = 10,
  model = "SIR",
  seed = 94128
)
```

## Probability sweeps

```{r probability-sweep}
simulate_inf_seq(
  input_matrix = initial,
  step = 0.1,
  model = "SIR",
  seed = 94128
)
```

## Heatmaps

`multiple_run_heatmap()` returns a matrix of infection counts by cell. Each
cell is counted once per run if it was infected at any point in that run,
including cells that later recovered, became susceptible again, or died. Set
`plot = TRUE` to draw the heatmap with base R graphics.

```{r heatmap}
heatmap_counts <- multiple_run_heatmap(
  prob_infect = 0.25,
  input_matrix = initial,
  runs = 10,
  model = "SIR",
  plot = FALSE,
  seed = 94128
)
heatmap_counts
```
