---
title: "SIR Simulation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{sirsimulation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(graphics)
library(SIRSimulation)
```

## Getting started

To get started with this package, you must first create a matrix that you can input into the simulation. You can use one of the methods to create a matrix in the package, or you can input a completely custom matrix. It is required that the input is a matrix data type, however, the size of the matrix does not matter.

You can create a matrix with a given chance of each cell starting as infected using the function.
```{r}
# Matrix with 25% chance of each cell starting infected
x <- create_matrix(5,5,0.25)
```
Similarly, you can use other designs such as starting the corners or very center cell as infected.
```{r}
# corners infected matrix
corners <- create_crnr_matrix(5,5)

# Center infected matrix
center <- create_cntr_matrix(15,15)
```
Or you can create your own custom matrix to use an your input.

## Running a simulation

To run a basic single simulation use the following function.
```{r}
# Run a basic simulation on matrix x
# Set probability of infection to 0.25
# Use default values auto_immunity FALSE and allow_death FALSE
# Since auto_immunity is FALSE, this will use the default immunity probability value of 0.5
simulate_sir(0.25,x)
```
However, there are some additional arguments we can add to the simulation call to increase the complexity of the simulation.

You can specify a specific probability of immunity by using the following arguments.
```{r}
# Run a basic simulation on matrix x
# Set probability of infection to 0.25
# Use default values auto_immunity FALSE and allow_death FALSE
# Set value for immunity probability to 0.34
simulate_sir(0.25,x,FALSE,0.34)
```

We can enable the auto_immunity parameter so that every cell that gets infected automatically gains immunity.
```{r}
# Run a basic simulation on matrix x
# Set probability of infection to 0.25
# Set auto_immunity to TRUE and use default value of allow_death FALSE
simulate_sir(0.25,x,TRUE)
```
Additionally, we can enable the allow_death parameter to enable cells to become Immune or Deceased after infection.
```{r}
# Run a basic simulation on matrix x
# Set probability of infection to 0.25
# Set auto_immunity to TRUE and immunity probability to 0.0
# Set allow_death to TRUE and fatality probability to 0.19
simulate_sir(0.25,x,TRUE,0.0,TRUE,0.19)
```
Lastly, we can disable the auto_immunity parameter and enable the allow_death parameter to enable cells to become Susceptible, Immune, or Deceased after infection.
```{r}
# Run a basic simulation on matrix x
# Set probability of infection to 0.25
# Set auto_immunity to FALSE and immunity probability to 0.35
# Set allow_death to TRUE and fatality probability to 0.12
simulate_sir(0.25,x,FALSE,0.35,TRUE,0.12)
```

## Simulating models over a sequence of infection probabilities

You can also run any of the above models over a sequence of different infection probabilities to compare results.
```{r}
# Run the sequence of infection probabilities using the basic simulation on matrix x from above
# Set step in sequence to 0.1
simulate_inf_seq(x,0.1)
```
You can modify the parameters of this function the same as the simulate_sir() function to modify the simulation settings.

## Simulate many model runs

You can run any of the above models a specified number times to get the average statistics.
```{r}
# Run a basic simulation on matrix x many times and report average statistics
# Set probability of infection to 0.25
# Set number of runs to 2
simulate_many_runs(0.25,x,2)
```
You can modify the parameters of this function the same as the simulate_sir() function to modify the simulation settings.

## Generate multiple runs heatmap

You can generate a heatmap of the results of multiple simulation runs to identify which cells are more likely to become infected.
```{r}
# Run a basic simulation on matrix x many times and generate heatmap
# Set probability of infection to 0.25
# Set number of runs to 2
multiple_run_heatmap(0.25,x,2)
```
You can modify the parameters of this function the same as the simulate_sir() function to modify the simulation settings.
